---
title: "DVEP_2_Analysis"
author: "Moura, G.S.P.L"
date: "`r format(Sys.Date(), '%d de %B de %Y')`"
output: html_document
---

```{r, eval= FALSE}
library(here)
library(tidyverse)
library(readxl)
library(lubridate)
library(stringr)
library(purrr)
library(gt)
library(jmv)
library(skimr)
```

# Modifying functions

## summarize_numerical_wider
Returns the dataframe in wide format
```{r}
summarize_numerical <- function(data, group_col = NULL, use_labels = TRUE) {
  # Extract variable labels, falling back to variable names if labels are missing
  variable_labels <- sapply(names(data), function(var) {
    label <- attr(data[[var]], "label")
    if (is.null(label) || !use_labels) var else label
  }, simplify = TRUE)
  
  # Identify numeric columns
  numeric_cols <- data %>% select(where(is.numeric)) %>% names()
  
  if (!is.null(group_col)) {
    # Grouped summary
    summary <- data %>%
      group_by(across(all_of(group_col))) %>%
      summarise(across(
        all_of(numeric_cols),
        ~ paste0(
          round(mean(.x, na.rm = TRUE), 1), 
          " (", 
          round(mean(.x, na.rm = TRUE) - 1.96 * sd(.x, na.rm = TRUE) / sqrt(sum(!is.na(.x))), 1), 
          "–", 
          round(mean(.x, na.rm = TRUE) + 1.96 * sd(.x, na.rm = TRUE) / sqrt(sum(!is.na(.x))), 1), 
          ")"
        ),
        .names = "{.col}"
      )) %>%
      pivot_longer(-all_of(group_col), names_to = "Variable", values_to = "Value") %>%
      mutate(Variable = variable_labels[Variable]) %>%
      pivot_wider(names_from = all_of(group_col), values_from = Value)  # Pivot to wide format
  } else {
    # Ungrouped summary
    summary <- data %>%
      summarise(across(
        all_of(numeric_cols),
        ~ paste0(
          round(mean(.x, na.rm = TRUE), 1), 
          " (", 
          round(mean(.x, na.rm = TRUE) - 1.96 * sd(.x, na.rm = TRUE) / sqrt(sum(!is.na(.x))), 1), 
          "–", 
          round(mean(.x, na.rm = TRUE) + 1.96 * sd(.x, na.rm = TRUE) / sqrt(sum(!is.na(.x))), 1), 
          ")"
        ),
        .names = "{.col}"
      )) %>%
      pivot_longer(everything(), names_to = "Variable", values_to = "Value") %>%
      mutate(Variable = variable_labels[Variable])
  }
  
  return(summary)
}

```

## compare_groups

Added an additional argument `return_df`, which, if set to TRUE, extracts the gt table into a data.frame

```{r}
compare_groups <- function(data, group_col = "allocation_group", use_labels = TRUE, return_df = FALSE) {
  
  # Extract variable labels, default to variable names if labels are missing
  variable_labels <- sapply(names(data), function(var) {
    label <- attr(data[[var]], "label") # Check for the 'label' attribute
    if (is.null(label) || !use_labels) var else label # Use variable name if label is missing
  }, simplify = TRUE)
  
  # Prepare results storage
  results <- data.frame(
    Variable = character(), # The variable being tested
    Test = character(),     # The type of test (t-test or chi-squared/Fisher's)
    Statistic = numeric(),  # The test statistic value
    P_value = numeric(),    # The p-value from the test
    stringsAsFactors = FALSE # Ensure strings are not converted to factors
  )
  
  # Loop through all columns in the data, excluding the grouping column
  for (var in setdiff(names(data), group_col)) {
    if (is.numeric(data[[var]])) {
      # Run t-test for numeric variables
      test_result <- t.test(data[[var]] ~ data[[group_col]])
      
      # Append the t-test results to the results data.frame
      results <- rbind(results, data.frame(
        Variable = variable_labels[var], # Use label or name
        Test = "t-test",                 
        Statistic = round(test_result$statistic, 2), 
        P_value = round(test_result$p.value, 4)
      ))
    } else if (is.factor(data[[var]]) || is.character(data[[var]])) {
      # Run chi-squared or Fisher's exact test
      contingency_table <- table(data[[var]], data[[group_col]])
      
      if (any(chisq.test(contingency_table)$expected < 5)) {
        # Use Fisher's exact test
        test_result <- fisher.test(contingency_table)
        results <- rbind(results, data.frame(
          Variable = variable_labels[var],
          Test = "Fisher's exact test",
          Statistic = NA,
          P_value = round(test_result$p.value, 4)
        ))
      } else {
        # Use chi-squared test
        test_result <- tryCatch(
          chisq.test(contingency_table), 
          error = function(e) list(statistic = NA, p.value = NA)
        )
        results <- rbind(results, data.frame(
          Variable = variable_labels[var],
          Test = "Chi-squared test",
          Statistic = round(test_result$statistic, 2),
          P_value = round(test_result$p.value, 4)
        ))
      }
    }
  }
  
  # If return_df = TRUE, return the results as a data frame
  if (return_df) {
    return(results)
  }
  
  # Return results as a gt table
  gt_table <- results %>%
    gt() %>%
    tab_header(
      title = "Hypothesis Test Results",
      subtitle = paste("Comparison of", group_col)
    ) %>%
    cols_label(
      Variable = "Variable",
      Test = "Test Type",
      Statistic = "Test Statistic",
      P_value = "P-value"
    )
  
  return(gt_table) # Return the formatted gt table
}

```


# Baseline
```{r}
file_path_baseline <- "Output/Baseline/Data_files"
```

```{r, eval=FALSE}
# Datas de inclusão e conclusão
summary(data$consent_date)

summary(data["conclusion_date"])

# Completaram a intervenção
summary(data_d1_exclusive$completed_intervention)
```


## Abandonos
Dos participantes incluídos, 23 (30.6%) não completaram os três meses de intervenção. A taxa de abandono foi maior nos homens, com 4 abandonos (40% dos homens incluídos), sendo 2 de cada grupo. No sexo feminino, a taxa de abandono foi de 29.2%, sendo oito do grupo placebo e 11 do grupo intervenção.

```{r, eval=FALSE}
# Abandonos
## Taxa de abandono
(23/75)*100 

## Abandono por sexo e grupo de alocação
data_d1_exclusive |> 
    filter(completed_intervention == "Não") |> 
    count(sex, allocation_group, name = "Total")

## Taxa de abandono por sexo
(4/10)*100 # Homens
(19/65)*100 # Mulheres

table(data$non_completion_reason)
```

### Contingency table by sex
```{r, eval=FALSE}
# Criar a tabela de contingência
tabela <- table(data_d1_exclusive$sex, data_d1_exclusive$completed_intervention)

# Exibir a tabela
print(tabela)

# Teste do qui-quadrado
teste_qui2 <- chisq.test(tabela)
teste_qui2$expected

# Exibir os resultados do teste do qui-quadrado
print(teste_qui2)
```

### Contingency table by sex, controlling for allocation group
```{r, eval=FALSE}
# Criar a tabela de contingência estratificada por allocation_group
tabela <- table(data_d1_exclusive$sex, 
                data_d1_exclusive$completed_intervention, 
                data_d1_exclusive$allocation_group)

# Exibir a tabela para verificar se está correta
print(tabela)

# Realizar o teste de Mantel-Haenszel
resultado_mh <- mantelhaen.test(tabela)

# Exibir o resultado do teste
print(resultado_mh)

```


### Log reg
```{r, eval=FALSE}
model <- glm(completed_intervention ~ sex * allocation_group, data = data_d1_exclusive, family = binomial)

summary(model)

exp(cbind(OR = coef(model), confint(model)))

```

## Participant's characteristics (data_d1_exclusive)
```{r, eval=FALSE}
# ALL
baseline_num <- summarize_numerical(data_d1_exclusive) # `summarize_numerical()`
baseline_num %>%
  gt() %>%
  tab_header(
    title = "Numerical Variable Summary",
    subtitle = "Overall Summary"
  )

baseline_cat <- summarize_categorical(data_d1_exclusive) # `summarize_categorical()`
baseline_cat %>%
  gt() %>%
  tab_header(
    title = "Categorical Variable Summary",
    subtitle = "Overall Summary"
  )

# BY GROUP
baseline_num_group <- summarize_numerical(data_d1_exclusive, "allocation_group") # `summarize_numerical()`
baseline_num_group %>%
  gt() %>%
  tab_header(
    title = "Numerical Variable Summary",
    subtitle = "Overall Summary"
  )

baseline_cat_group <- summarize_categorical(data_d1_exclusive, "allocation_group") # `summarize_categorical()`
baseline_cat_group %>%
  gt() %>%
  tab_header(
    title = "Categorical Variable Summary",
    subtitle = "Overall Summary"
  )

# COMPARE GROUPS
group_comparison <- compare_groups(data_d1_exclusive) #`compare_groups`
print(group_comparison)
```

Descriptive:
Drugs that might induce weight loss
```{r, eval=FALSE}
drugs_w_loss_descriptive <- I22_drugs_R |> 
    dplyr::filter(stringr::str_detect(drugs_sql, 
        "C61939|C62012|C506_1|C1278_2|C1278_1|C1278_3|C47764_1|C47764_2|C61680"
    )) |> 
    dplyr::pull(descriptive) |> 
    stringr::str_extract("^[^0-9]+") |> 
    stringr::str_trim() |> 
    base::unique() |> 
    cat(sep = ", ")

drugs_w_gain <- I22_drugs_R |> 
    filter(str_detect(drugs_sql, 
        "C61879|C62005|C61917_2|C29416|C29536_2")) |> 
    dplyr::pull(descriptive) |> 
    stringr::str_extract("^[^0-9]+") |> 
    stringr::str_trim() |> 
    base::unique() |> 
    cat(sep = ", ")
```


**Clear** Participant's characteristics (data_d1_exclusive)
```{r, eval=FALSE}
# CLEAR
rm(baseline_num)
rm(baseline_cat)
rm(baseline_num_group)
rm(baseline_cat_group)
rm(group_comparison)
rm(drugs_w_loss_descriptive)

```

Comparação da proporção do sexo dos participantes segundo o grupo de alocação
Na verdade essa análise foi redundante - já foi feita em `baseline_cat_group`
```{r, eval=FALSE}
group_by_sex <- data_d1_exclusive |> 
    group_by(sex, allocation_group) |> 
    count() |> 
    pivot_wider(names_from = allocation_group, values_from = n, values_fill = 0) |>
    column_to_rownames(var = "sex") |> # Converts the 'sex' column into row names
    as.matrix()

group_by_sex_chisq <- chisq.test(group_by_sex)
group_by_sex_chisq
group_by_sex_FET <- fisher.test(group_by_sex)
group_by_sex_FET

rm(group_by_sex)
rm(group_by_sex_chisq)
rm(group_by_sex_FET)
```


## Lab exames
```{r, eval=FALSE}
# filter first visit and relevant variables
labs_baseline <- I27_labs_R |> 
    filter(
        visit == 1
    ) |> 
    select(
        -visit, -labs_date, -labs_beta_hcg , -labs_notes
    ) |> left_join(
        data_d1_exclusive |> 
            select(record_id, allocation_group),
                   by = "record_id"
                  )

baseline_labs_baseline <- summarize_numerical(labs_baseline) #all
baseline_labs_baseline_groups <- summarize_numerical(labs_baseline,"allocation_group") #groups
baseline_labs_baseline_groups_compare <- compare_groups(labs_baseline) #compare
print(baseline_labs_baseline_groups_compare) #print
```

**Clear** Lab exams
```{r, eval=FALSE}
rm(labs_baseline)
rm(baseline_labs_baseline)
rm(baseline_labs_baseline_groups)
rm(baseline_labs_baseline_groups_compare)
```

**CPR**
```{r, eval=FALSE}

plot(labs_baseline$labs_crp)

cpr_bp <- boxplot(labs_baseline$labs_crp_log)
cpr_bp$out

par(mfrow = c(1, 2)) 
hist(labs_baseline$labs_crp)
hist(labs_baseline$labs_crp_log)

labs_baseline$labs_crp_log <- log(labs_baseline$labs_crp)

summary(labs_baseline$labs_crp_log)

plot(labs_baseline$labs_crp_log, data_filtered$phaseangle[data_filtered$visit == 1])


skim(labs_baseline)


rm(cpr_bp)
rm(labs_baseline$labs_crp_log)
```


## BIA
### Baseline
```{r, eval=FALSE}
bia_baseline <- data_bia |>
    filter(
        visit == 1
    ) |> 
    select(
        record_id, phaseangle:w_ecw
        ) |> 
    dplyr::left_join(
        data_d1_exclusive |> 
            select(record_id, allocation_group, sex),
                   by = "record_id"
        )
```

### Comparing between males and females
```{r, eval=FALSE}
#by sex
bia_baseline_sex <- summarize_numerical(bia_baseline, "sex") 
bia_baseline_sex_compare <- compare_groups(bia_baseline,"sex") # compare
print(bia_baseline_sex_compare) # print
```

### Males
```{r, eval=FALSE}
# Filter males
BIA_V1_M <- bia_baseline |> 
    filter(
        sex == "Masculino"
    )

#summarize_numerical()
BIA_V1_M_SN <- summarize_numerical(BIA_V1_M, "allocation_group",use_labels = TRUE)

# compare_groups()
BIA_V1_M_C <- compare_groups(BIA_V1_M, group_col = "allocation_group",use_labels = TRUE, return_df = TRUE) 

# join tables
bia_males <- BIA_V1_M_SN %>%
    left_join(y = BIA_V1_M_C %>% 
                  select(Variable, P_value),
              by = join_by(Variable)
    ) %>%
    mutate(
        P_value = round(P_value,3)
    ) %>% 
    filter(
        Variable != "Nome do arquivo"
    ) %>% 
    rename(
         Parâmetro = Variable,
         `Grupo placebo [média (IC 95%)]` = `Grupo A`,
         `Grupo Eclipta [média (IC 95%)]` = `Grupo B`,
         `Valor p [Teste t]` = P_value
    ) %>% 
    mutate(Parâmetro = recode(Parâmetro, 
                            "Altura" = "Altura (m)", 
                            "Peso" = "Peso (kg)", 
                            "Índice de Massa Corporal" = "Índice de Massa Corporal (kg/m²)", 
                            "Cintura" = "Circunferência Abdominal (m)", 
                            "Nível de Atividade Física" = "Nível de Atividade Física (PAL)", 
                            "Ângulo de Fase" = "Ângulo de Fase (°)", 
                            "Média de Resistência" = "Resistência (Ω)", 
                            "Média de Reatância" = "Reatância (Ω)", 
                            "Massa de Gordura Absoluta" = "Massa gorda Absoluta (kg)", 
                            "Índice de Massa de Gordura" = "Índice de Massa gorda (kg/m²)", 
                            "Massa de Gordura Relativa" = "Massa gorda Relativa (%)", 
                            "Tecido Adiposo Visceral" = "Tecido Adiposo Visceral (kg)", 
                            "Massa Livre de Gordura Absoluta" = "Massa Livre de Gordura Absoluta (kg)", 
                            "Índice de Massa Livre de Gordura" = "Índice de Massa Livre de Gordura (kg/m²)", 
                            "Massa Livre de Gordura Relativa" = "Massa Livre de Gordura Relativa (%)", 
                            "Massa Muscular Esquelética do Corpo" = "MME Corporal Total (kg)", 
                            "Massa Muscular Esquelética do Tronco" = "MME do Tronco (kg)", 
                            "Massa Muscular Esquelética do Braço Direito" = "MME do Braço Direito (kg)", 
                            "Massa Muscular Esquelética do Braço Esquerdo" = "MME do Braço Esquerdo (kg)", 
                            "Massa Muscular Esquelética da Perna Direita" = "MME da Perna Direita (kg)", 
                            "Massa Muscular Esquelética da Perna Esquerda" = "MME da Perna Esquerda (kg)", 
                            "Água Corporal Total" = "Água Corporal Total (L)", 
                            "Água Extracelular" = "Água Extracelular (L)", 
                            "Gasto Energético Total" = "Gasto Energético Total (kcal)", 
                            "Gasto Energético em Repouso" = "Gasto Energético em Repouso (kcal)", 
                            "Energia Armazenada" = "Energia Armazenada (kcal)")
        ) %>%
    mutate(Parâmetro = factor(Parâmetro, levels = c(
    "Altura (m)", "Peso (kg)", "Índice de Massa Corporal (kg/m²)", "Circunferência Abdominal (m)", 
    "Nível de Atividade Física (PAL)", "Ângulo de Fase (°)", "Resistência (Ω)", "Reatância (Ω)", 
    "Massa gorda Absoluta (kg)", "Índice de Massa gorda (kg/m²)", "Massa gorda Relativa (%)", 
    "Tecido Adiposo Visceral (kg)", "Massa Livre de Gordura Absoluta (kg)", 
    "Índice de Massa Livre de Gordura (kg/m²)", "Massa Livre de Gordura Relativa (%)", 
    "MME Corporal Total (kg)", "MME do Tronco (kg)", "MME do Braço Direito (kg)", 
    "MME do Braço Esquerdo (kg)", "MME da Perna Direita (kg)", "MME da Perna Esquerda (kg)", 
    "Água Corporal Total (L)", "Água Extracelular (L)", "Gasto Energético Total (kcal)", 
    "Gasto Energético em Repouso (kcal)", "Energia Armazenada (kcal)"))
    ) %>% 
    arrange(Parâmetro)

write_csv(bia_males, here("Output", "Baseline", "Tables", "bia_males.csv"))
```

### Females
```{r, eval=FALSE}
# Filter males
BIA_V1_F <- bia_baseline |> 
    filter(
        sex == "Feminino"
    )

#summarize_numerical()
BIA_V1_F_SN <- summarize_numerical(BIA_V1_F, "allocation_group",use_labels = TRUE)

# compare_groups()
BIA_V1_F_C <- compare_groups(BIA_V1_F, group_col = "allocation_group",use_labels = TRUE, return_df = TRUE) 

# join tables
bia_females <- BIA_V1_F_SN %>%
    left_join(y = BIA_V1_F_C %>% 
                  select(Variable, P_value),
              by = join_by(Variable)
    ) %>%
    mutate(
        P_value = round(P_value,3)
    ) %>% 
    filter(
        Variable != "Nome do arquivo"
    ) %>% 
    rename(
         Parâmetro = Variable,
         `Grupo placebo [média (IC 95%)]` = `Grupo A`,
         `Grupo Eclipta [média (IC 95%)]` = `Grupo B`,
         `Valor p [Teste t]` = P_value
    ) %>% 
    mutate(Parâmetro = recode(Parâmetro, 
                            "Altura" = "Altura (m)", 
                            "Peso" = "Peso (kg)", 
                            "Índice de Massa Corporal" = "Índice de Massa Corporal (kg/m²)", 
                            "Cintura" = "Circunferência Abdominal (m)", 
                            "Nível de Atividade Física" = "Nível de Atividade Física (PAL)", 
                            "Ângulo de Fase" = "Ângulo de Fase (°)", 
                            "Média de Resistência" = "Resistência (Ω)", 
                            "Média de Reatância" = "Reatância (Ω)", 
                            "Massa de Gordura Absoluta" = "Massa gorda Absoluta (kg)", 
                            "Índice de Massa de Gordura" = "Índice de Massa gorda (kg/m²)", 
                            "Massa de Gordura Relativa" = "Massa gorda Relativa (%)", 
                            "Tecido Adiposo Visceral" = "Tecido Adiposo Visceral (kg)", 
                            "Massa Livre de Gordura Absoluta" = "Massa Livre de Gordura Absoluta (kg)", 
                            "Índice de Massa Livre de Gordura" = "Índice de Massa Livre de Gordura (kg/m²)", 
                            "Massa Livre de Gordura Relativa" = "Massa Livre de Gordura Relativa (%)", 
                            "Massa Muscular Esquelética do Corpo" = "MME Corporal Total (kg)", 
                            "Massa Muscular Esquelética do Tronco" = "MME do Tronco (kg)", 
                            "Massa Muscular Esquelética do Braço Direito" = "MME do Braço Direito (kg)", 
                            "Massa Muscular Esquelética do Braço Esquerdo" = "MME do Braço Esquerdo (kg)", 
                            "Massa Muscular Esquelética da Perna Direita" = "MME da Perna Direita (kg)", 
                            "Massa Muscular Esquelética da Perna Esquerda" = "MME da Perna Esquerda (kg)", 
                            "Água Corporal Total" = "Água Corporal Total (L)", 
                            "Água Extracelular" = "Água Extracelular (L)", 
                            "Gasto Energético Total" = "Gasto Energético Total (kcal)", 
                            "Gasto Energético em Repouso" = "Gasto Energético em Repouso (kcal)", 
                            "Energia Armazenada" = "Energia Armazenada (kcal)")
        ) %>%
    mutate(Parâmetro = factor(Parâmetro, levels = c(
    "Altura (m)", "Peso (kg)", "Índice de Massa Corporal (kg/m²)", "Circunferência Abdominal (m)", 
    "Nível de Atividade Física (PAL)", "Ângulo de Fase (°)", "Resistência (Ω)", "Reatância (Ω)", 
    "Massa gorda Absoluta (kg)", "Índice de Massa gorda (kg/m²)", "Massa gorda Relativa (%)", 
    "Tecido Adiposo Visceral (kg)", "Massa Livre de Gordura Absoluta (kg)", 
    "Índice de Massa Livre de Gordura (kg/m²)", "Massa Livre de Gordura Relativa (%)", 
    "MME Corporal Total (kg)", "MME do Tronco (kg)", "MME do Braço Direito (kg)", 
    "MME do Braço Esquerdo (kg)", "MME da Perna Direita (kg)", "MME da Perna Esquerda (kg)", 
    "Água Corporal Total (L)", "Água Extracelular (L)", "Gasto Energético Total (kcal)", 
    "Gasto Energético em Repouso (kcal)", "Energia Armazenada (kcal)"))
    ) %>% 
    arrange(Parâmetro)

write_csv(bia_females, here("Output", "Baseline", "Tables", "bia_females.csv"))
```

## Handgrip
```{r, eval=FALSE}
handgrip <- data |> 
    select(record_id, event_name, handgrip_right_mean, handgrip_left_mean) |> 
    filter(
        event_name == "1visit_arm_1"
    ) |> 
    filter(
        !is.na(handgrip_right_mean) | !is.na(handgrip_left_mean)
    ) |> 
    left_join(
        data_d1_exclusive |> 
            select(
                record_id, sex, allocation_group
            ),
        by = "record_id"
    ) |> 
    mutate(
        handgrip = rowMeans(
            across(c(handgrip_left_mean, handgrip_right_mean)),
            na.rm = TRUE
        )
    )
#write_excel_csv(handgrip,paste0(file_path_baseline,"handgrip.csv"))

summarize_numerical(handgrip,"sex")
compare_groups(handgrip,"sex")

# handgrip |> 
    filter(
        sex == "Masculino"
    ) |> 
    group_by(
        allocation_group
    ) |> 
    summarize(
        `Média da mão direita` = mean(handgrip_right_mean, na.rm = TRUE),
        `Média da mão esquerda` = mean(handgrip_left_mean, na.rm = TRUE),
        `Média das mãos` = mean(handgrip, na.rm = TRUE),
    )

# Males
handgrip_man <- handgrip |> 
    filter(
        sex == "Masculino"
    )
handgrip_man_num <- print(summarize_numerical(handgrip_man, "allocation_group"))
handgrip_man_compare <- print(compare_groups(handgrip_man, "allocation_group"))

# Females
handgrip_fem <- handgrip |> 
    filter(
        sex == "Feminino"
    )
handgrip_fem_num <- print(summarize_numerical(handgrip_fem, "allocation_group"))
handgrip_fem_compare <- print(compare_groups(handgrip_fem, "allocation_group"))
```



## Questionnaires
```{r, eval=FALSE}
questionnaires <- data_filtered |> 
    ungroup() |> 
    filter(
        visit == 1
    ) |> 
    select(
        record_id, allocation_group, sex, whoqol_score_overall:ecap_score
    )

# write_excel_csv(questionnaires, paste0(file_path_baseline,"questionnaires.csv"))

questionnaires_num <- summarize_numerical(questionnaires)
questionnaires_num_groups <- summarize_numerical(questionnaires,"allocation_group")
questionnaires_num_groups_compare <- compare_groups(questionnaires,"allocation_group")

# By sex
questionnaires_sex <- summarize_numerical(questionnaires,"sex")
questionnaires_sex_compare <- compare_groups(questionnaires,"sex")
print(questionnaires_sex_compare)

# By allocation_group
questionnaires_num <- summarize_numerical(questionnaires,"allocation_group")
questionnaires_num_compare <- compare_groups(questionnaires,"allocation_group")
print(questionnaires_sex_compare)

# ANOVA to search for interaction between sex and allocation
anova_whoqol <- aov(whoqol_score_overall ~ sex * allocation_group, data = questionnaires)
summary(anova_whoqol)
anova_depression <- aov(dass_score_depression ~ sex * allocation_group, data = questionnaires)
summary(anova_depression)
anova_anxiety <- aov(dass_score_anxiety ~ sex * allocation_group, data = questionnaires)
summary(anova_anxiety)
anova_stress <- aov(dass_score_stress ~ sex * allocation_group, data = questionnaires)
summary(anova_stress)
anova_ecap <- aov(ecap_score ~ sex * allocation_group, data = questionnaires)
summary(anova_ecap)

```
# Adverse Events
```{r, eval=FALSE}
adverse_events <- read_csv('/Users/gustavosplmoura/Library/Mobile Documents/com~apple~CloudDocs/Medicina/Biblioteca/Research/Data Science/Data Science/PROJECTS/DVEP/Analysis/adverse_events.csv')

adverse_events$`Nome completo` <- substr(adverse_events$`Nome completo`,1,2)
colnames(adverse_events) <- trimws(colnames(adverse_events))

adverse_events_yes <- adverse_events |>
    filter(
        `Algum Evento Adverso neste ciclo?` == "Sim"
    ) |> 
    mutate(
        `Nome completo` = as.integer(`Nome completo`)
    ) |>
    left_join(
        data_d1_exclusive |>
            select(record_id, sex, allocation_group),
        by = join_by(`Nome completo` == record_id)
    ) |> 
    mutate(
        `Event Name` = as.factor(`Event Name`),
        `Evento Adverso Grave?  ` = as.factor(`Evento Adverso Grave?  `),
        Grau = as.factor(Grau),
        Classificação = as.factor(Classificação),
        `Atribuição de causalidade` = as.factor(`Atribuição de causalidade`),
        `Ação Realizada` = as.factor(`Ação Realizada`),
        Seguimento = as.factor(Seguimento),
        `Nome completo` = as.character(`Nome completo`)
    )

#summarize_categorical
adverse_events_cat <- summarize_categorical(adverse_events_yes)
## write_excel_csv(adverse_events_cat, paste0(file_path_baseline, "adverse_events_cat.csv"))

#baseR: summary
summary(adverse_events_yes)
```

**Association between visit (time point) and severity of adverse event**
Testing for an association between visit (time point) and severity of adverse events to evaluate whether the likelihood of experiencing severe adverse events changes as the intervention progresses (Chi-squared test of independence)
```{r, eval=FALSE}
# contingency table
table_ae <- table(adverse_events_yes$`Event Name`, adverse_events_yes$Grau)
print(table_ae)

# Chi-square
chisq_test <- chisq.test(table_ae)
print(chisq_test)
chisq_test$expected
## Many cells with expected counts < 5 -> FET would be more appropriate
fisher_test <- fisher.test(table_ae)
print(fisher_test)

adverse_events_yes$allocation_group

```

--> Running Mantel-Haenszel test to control for the intervention group
```{r, eval=FALSE}
library(gmodels)

# CrossTable to explore the stratified data
CrossTable(
  x = adverse_events_yes$`Event Name`, 
  y = adverse_events_yes$Grau, 
  strata = adverse_events_yes$allocation_group,
  prop.chisq = FALSE, 
  prop.r = FALSE, 
  prop.c = FALSE, 
  prop.t = FALSE
)

# Create a 3D contingency table stratified by allocation group
table_visit_severity_allocation <- table(
    adverse_events_yes$`Event Name`, adverse_events_yes$Grau, adverse_events_yes$allocation_group
    )
print(table_visit_severity_allocation)

# Mantel-Haenszel test
mantel_test <- mantelhaen.test(table_visit_severity_allocation)
print(mantel_test)

```




**By severity**
```{r, eval=FALSE}
table_ae <- table(adverse_events_yes$Classificação, adverse_events_yes$Grau)
print(table_ae)
```
**By causality**
```{r, eval=FALSE}
table_ae <- table(adverse_events_yes$Classificação, adverse_events_yes$`Atribuição de causalidade`)
View(table_ae)
```

**Gastrointestinal**
```{r, eval=FALSE}
ae_gastrointestinal <- adverse_events_yes |> 
    filter(
        Classificação == "Gastrointestinal"
    )
```

**Grouping events by classification**
```{r, eval=FALSE}

# Group by Classificação and Evento Adverso, count occurrences, and calculate percentages
summary_data <- adverse_events_yes %>%
  group_by(Classificação, `Evento Adverso`) %>%
  summarise(Count = n(), .groups = "drop") %>%
  mutate(Percentage = (Count / sum(Count)) * 100) %>%
  arrange(Classificação, desc(Count))

# Create the formatted output
formatted_output <- summary_data %>%
  group_by(Classificação) %>%
  summarise(
    Events = paste0(`Evento Adverso`, " (", round(Percentage, 1), "%)", collapse = ", "),
    .groups = "drop"
  ) %>%
  mutate(Result = paste0(Classificação, ": ", Events))

# Print the result
cat(paste(formatted_output$Result, collapse = "\n"))

# Save file
write_lines(paste(formatted_output$Result, collapse = "\n"), paste0(file_path_baseline, "eventos_adversos_por_sistema.txt"))


```


**Clear**
```{r, eval=FALSE}
rm(questionnaires)
rm(questionnaires_num)
rm(questionnaires_compare)

```


**Clear**
```{r, eval=FALSE}
rm()
rm()
rm()
rm()
rm()
```


**Clear**
```{r, eval=FALSE}
rm()
rm()
rm()
rm()
rm()
```


**Clear**
```{r, eval=FALSE}
rm()
rm()
rm()
rm()
rm()
```



