---
title: "DVEP Data Analysis"
project: "Effect of Eclipta prostrata (L.) L. (Asteraceae) on bioelectrical impedance phase angle in adults with grade I obesity (DVEP)"
PID: "REDCap 1958"
author: "Gustavo Santos Paiva Laender Moura"
date: "2024-12-16"
output: html_document
---

# FIRST STEPS

## Exporting data from REDCap
1. PID 1958
2. Data Exports, Reports, and Stats
3. Choose "Export Data" from option "A - All data (all records and fields)"
4. Export format: CSV (raw data)
5. De-identification options: "Remove all identifier fields" YES
6. Additional export options: "Export survey identifier field and survey timestamp field(s)?" YES
7. Advanced data formatting options
- "Export gray Form Status fields with blank value"
- comma as default separator
- Use period/full stop (.) as decimal
8. Export file to "/Users/gustavosplmoura/Library/Mobile Documents/com~apple~CloudDocs/Medicina/Biblioteca/Research/Data Science/Data Science/PROJECTS/DVEP"
9. Rename file to "data_dvep.csv"
Note: resulting file must have 780 variables (last variable on column ACZ: "anexos_complete")

## Getting Started with R
1. Loading libraries
2. Reading CSV data files (Tidyverse)
3. Remove identifying data from record_id
4. Renaming variables
5. Assigning correct variable types

```{r, results = 'hide'}
# Clear existing data and graphics
rm(list = ls())
graphics.off()
cat("\014")  # Clear any pending RStudio sessions or temporary files

# Load necessary libraries
library(Hmisc)
library(tidyverse)

# Set working directory
setwd("/Users/gustavosplmoura/Library/Mobile Documents/com~apple~CloudDocs/Medicina/Biblioteca/Research/Data Science/Data Science/PROJECTS/DVEP")

# 2. Reading CSV data files (Tidyverse)
data  <- read_csv(
  "data_dvep.csv",
  col_names = TRUE,
  col_types = NULL,
  col_select = NULL,
  id = NULL,
  locale = default_locale(),
  na = c("", "NA", "NI", "UNK", "NASK", "ASKU", "INV"),
  quote = "\"",
  comment = "",
  trim_ws = TRUE,
  skip = 0, # Number of lines to skip before reading data
  n_max = Inf, # Maximum number of lines to read.
  guess_max = 1000,
  name_repair = "unique",
  num_threads = readr_threads(),
  progress = show_progress(),
  show_col_types = TRUE,
  skip_empty_rows = TRUE,
  lazy = should_read_lazy()
)

data_codebook <- read_csv(
  "data_codebook.csv",
  col_names = TRUE,
  col_types = NULL,
  col_select = NULL,
  id = NULL,
  locale = default_locale(),
  na = c("", "NA", "NI", "UNK", "NASK", "ASKU", "INV"),
  quote = "\"",
  comment = "",
  trim_ws = TRUE,
  skip = 0, # Number of lines to skip before reading data
  n_max = Inf, # Maximum number of lines to read.
  guess_max = 1000,
  name_repair = "unique",
  num_threads = readr_threads(),
  progress = show_progress(),
  show_col_types = TRUE,
  skip_empty_rows = TRUE,
  lazy = should_read_lazy()
)
data_structure  <- read_csv(
  "data_structure.csv",
  col_names = TRUE)

data_NCIT  <- read_csv(
  "data_NCIT.csv",
  col_names = TRUE)

# 3. Remove identifying data from record_id
data$record_id <- substr(data$record_id,1,2)

# 4. Renaming variables
rename_vector <- setNames(object = colnames(data), data_codebook$variable)
data <- data |>
    rename(!!!rename_vector)

rm(rename_vector)

# 5. Assingning correct variable types
convert_col_type <- function(column, type) {
  switch(type,
         c = as.character(column),
         d = as.numeric(column),
         i = as.integer(column),
         l = as.logical(column),
         D = as.Date(column, format = "%Y-%m-%d"),
         T = as.POSIXct(column, format = "%Y-%m-%d %H:%M:%S"),
         f = as.factor(column),
         column  # Default (no change)
  )
}

data <- data |>
    mutate(
        across( # apply a function to multiple columns at once
            .cols = # .cols: Specifies which columns you want to change
                all_of(data_codebook$variable), 
            .fns = # .fns: Defines the function to apply to each column
                ~ convert_col_type(.x, data_codebook$col_types[which(data_codebook$variable == cur_column())])
))
rm(convert_col_type)
```

### Assing correct variable types
1. Excel formula to convert REDCap field types to tidyverse tibble nomenclature
    =IF(
    J2="calc","d",IF(
    J2="radio","f",IF(
    J2="notes","c",IF(
    J2="dropdown","f",IF(
    J2="slider","i",if(
        and(J2="text",N2=""),"c",if(
        and(J2="text",N2="date_dmy"),"D",if(
        and(J2="text",N2="integer"),"i",if(
        and(J2="text",left(N2,6)="number"),"d",if(
        and(J2="text",N2="datetime_dmy"),"T",if(
        and(J2="text",N2="time_mm_ss"),"t",if(
        and(J2="text",N2="time"),"t",""
        ))))))))))))

### Checking each column's variable type
If you are using dplyr, you can use glimpse(), which is like str() but optimized for tibbles. It shows a cleaner, more readable summary.
<chr>: Character
<dbl>: Double (floating-point number)
<int>: Integer
<fct>: Factor
<date>: Date
<dttm>: Date-Time (POSIXct)

sapply() (Show all column types)
sapply(data, class)

# TIDYVERSE

## Create one tibble for each instrument
Steps:
1. Create list of REDCap instruments and associated variables
2. Based on the instruments' list, create one tibble for each instrument 

Coding:
1. Create list of REDCap instruments and associated variables
```{r}
instruments <- list(
    I01_elegibility = c("record_id", "event_name", "date_birth", "age", "ubs", "contact_options", "contact_hours", "contact_days", "transport_research_center", "research_source_info", "availability_comments_yn", "availability_comments", "sex", "pregnant_nursing_yn", "eleg_height", "eleg_weight", "bmi_reported", "eleg_comorbidity_yn", "comorbidity_list", "drug_use_yn", "drug_list", "high_risk_pregnancy_ineffective_contraceptives", "high_risk_pregnancy_inconsistent_contraceptives", "high_risk_pregnancy_unprotected_sex", "high_risk_pregnancy_infertility_treatment", "high_risk_pregnancy_postpartum", "high_risk_pregnancy_none", "high_risk_fem_noinfo", "high_risk_fem_unknown", "high_risk_fem_notasked", "high_risk_fem_askunknown", "high_risk_fem_invalid", "high_risk_fem_na", "pregnancy_test_and_contraceptive_agreement", "low_risk_pregnancy_effective_contraceptives", "menopause", "low_risk_pregnancy_surgical_sterilization", "low_risk_pregnancy_no_sex_6_months", "low_risk_pregnancy_no_sex", "low_risk_pregnancy_infertility_diagnosis", "low_risk_pregnancy_exclusive_homo_sexual_behavior", "low_risk_fem_noinfo", "low_risk_fem_unknown", "low_risk_fem_notasked", "low_risk_fem_askunknown", "low_risk_fem_invalid", "low_risk_fem_na", "contraceptive_continuation_agreement", "lab_notes", "consultation_schedule_date", "eleg_lab_request_date", "lab_schedule_date", "eleg_lab_location", "lab_collected_yn", "lab_checked_yn", "lab_finding_yn", "lab_exclusion_yn", "bhcg", "contraception", "preserved_hormone", "contraception_ready_yn", "contraception_ineligibility", "consent_sent_yn", "consent_read_yn", "consent_questions_yn", "consent_questions_2_yn", "consent_questions_cleared_yn", "participant_desire_yn", "first_visit_date", "sms_consent_yn", "block_alerts_1st_yn", "interview_datetime", "eleg_interviewer", "eleg_complete"),
    I02_tcle = c("record_id","event_name", "consent_signed_yn", "consent_date", "consent_upload", "tcle_complete"),
    I03_demographic = c("record_id","event_name", "race", "race_other", "marital_status", "education_years", "employment_status", "household_size", "income_level", "demographic_metadata", "demographic_complete"),
    I04_whoqol = c("record_id","event_name", "whoqol_timestamp", "whoqol_1_quality", "whoqol_2_health", "whoqol_3_pain", "whoqol_4_treatment", "whoqol_5_enjoyment", "whoqol_6_meaning", "whoqol_7_concentration", "whoqol_8_security", "whoqol_9_environment", "whoqol_10_energy", "whoqol_11_appearance", "whoqol_12_finances", "whoqol_13_information", "whoqol_14_leisure", "whoqol_15_mobility", "whoqol_16_sleep", "whoqol_17_activities", "whoqol_18_work", "whoqol_19_selfesteem", "whoqol_20_relationships", "whoqol_21_sexual", "whoqol_22_support", "whoqol_23_housing", "whoqol_24_health_services", "whoqol_25_transport", "whoqol_26_negativity", "whoqol_needed_help", "whoqol_comment", "whoqol_score_overall", "whoqol_score_health", "whoqol_score_physical", "whoqol_score_psychological", "whoqol_score_social", "whoqol_score_environment", "pdf_upload_wb", "whoqol_complete"),
    I05_dass = c("record_id","event_name", "dass_timestamp", "dass_1_not_calm", "dass_2_drymouth", "dass_3_not_positive", "dass_4_hard_breath", "dass_5_no_initiative", "dass_6_exaggeration", "dass_7_tremor", "dass_8_nervous", "dass_9_worry", "dass_10_no_desire", "dass_11_agitation", "dass_12_not_relaxed", "dass_13_depression", "dass_14_intolerance", "dass_15_panic", "dass_16_no_enthusiasm", "dass_17_no_selfworth", "dass_18_too_emotional", "dass_19_palpitation", "dass_20_fear", "dass_21_no_meaning", "dass_score_depression", "dass_score_anxiety", "dass_score_stress", "dass21_upload", "dass_complete"),
    I06_ecap = c("record_id","event_name", "ecap_timestamp", "ecap1", "ecap2", "ecap3", "ecap4", "ecap5", "ecap6", "ecap7", "ecap8", "ecap9", "ecap10", "ecap11", "ecap12", "ecap13", "ecap14", "ecap15", "ecap16", "ecap_response_date", "ecap_1_body_image", "ecap_2_eating_speed", "ecap_3_impulse", "ecap_4_emotional_eating", "ecap_5_hunger", "ecap_6_guilt", "ecap_7_diet_failure", "ecap_8_binge_quantity", "ecap_9_compensate", "ecap_10_control", "ecap_11_fullness", "ecap_12_social_eating", "ecap_13_meal_frequency", "ecap_14_worry", "ecap_15_food_thoughts", "ecap_16_hunger_awareness", "ecap_score", "ecap_pdf_upload", "ecap_complete"),
    I07_bodycount = c("record_id","event_name", "height", "height_checked_cm", "weight", "weight_checked_kg", "abdomen", "abdominal_circ_checked_cm", "arm", "arm_circ_checked_cm", "anthro_check_error", "anthro_check", "bmi", "weight_maxideal", "weight_adjusted", "kcal_kg20", "kcal_kg22", "diet_energy", "weight_loss_kg", "weight_loss_percent", "abdomen_loss_kg", "abdomen_loss_percent", "bmi_loss_absolute", "bmi_loss_percent", "bodycount_complete"),
    I08_bp_limb = c("record_id","event_name", "bpbia_form_open_date", "bpbia_examiner", "bpbia_r_sw", "cuff_size", "reference_systolic_right", "reference_diastolic_right", "reference_mean_right", "reference_systolic_left", "reference_diastolic_left", "reference_mean_left", "reference_member", "bp_limb_complete"),
    I09_bp = c("record_id","event_name", "bp_form_open_date", "bp_examiner", "bp_start_sw", "heart_rate", "oxigen_saturation", "systolic_1", "diastolic_1", "bp_mean_1", "bp_first_measurement_sw", "systolic_2", "bp_r21s_checked", "diastolic_2", "bp_r22d_checked", "bp_mean_2", "bp_second_measurement_sw", "systolic_3", "bp_r31s_checked", "diastolic_3", "bp_r32d_checked", "bp_mean_3", "bp_sys_diff", "bp_third_measurement_sw", "systolic_4", "bp_r41s_checked", "diastolic_4", "bp_r42d_checked", "bp_mean_4", "systolic_mean", "systolic_stdev", "diastolic_mean", "diastolic_stdev", "mean_bp_mean", "mean_bp_stdev", "bp_show_metadata", "bp_complete"),
    I10_bia = c("record_id","event_name", "bia_raw_files", "bia_form_open_date", "bia_examiner", "smoked_24h", "exercised_24h", "alcohol_24h", "room_temperature", "axillary_temperature", "light_clothes", "removed_objects", "hairy_limbs", "removed_hair", "cleaned_skin", "last_meal_time", "last_liquid_time", "bed_rest_10min", "bia_timestamp", "time_fasted_food", "time_fasted_liquid", "resistance", "resistance_checked", "reactance", "reactance_checked", "phase_angle", "phase_angle_checked", "bia_show_metadata", "bia_complete"),
    I11_handgrip = c("record_id","event_name", "grip_limitations_yn", "grip_rh_limitations", "grip_lh_no_limitations", "grip_lh_paralysis", "grip_lh_cast", "grip_lh_bandaged", "grip_lh_missing_thumb", "grip_lh_other_limitations", "grip_lh_no_info", "grip_lh_unknown", "grip_lh_notasked", "grip_lh_askunknown", "grip_lh_invalid", "grip_lh_na", "grip_recent_surgery_yn", "grip_exclude_test_yn", "grip_surgery_history_yn", "grip_dominance", "grip_pain_last7days_yn", "grip_rhpain", "grip_lhpain", "grip_test_procedure_understanding_yn", "handgrip_right_1", "grip_rh_validation1", "grip_rh_stopwatch1", "handgrip_left_1", "grip_lh_validation1", "grip_lh_stopwatch1", "handgrip_right_2", "grip_rh_validation2", "grip_rh_stopwatch2", "handgrip_left_2", "grip_lh_validation2", "grip_lh_stopwatch2", "handgrip_right_3", "grip_rh_validation3", "handgrip_left_3", "grip_lh_validation3", "handgrip_right_mean", "handgrip_right_stdev", "handgrip_left_mean", "handgrip_left_stdev", "grip_see_metadata", "handgrip_complete"),
    I12_nutrition = c("record_id","event_name", "frequency_bowel", "frequency_solid_stool", "frequency_diarrhea", "contipation_yn", "frequency_laxative", "urinary_frequency", "water_consumption", "hair_loss", "nutrition_complete"),
    I13_food_allergy = c("record_id","event_name", "food_alergy", "peanut_allergy", "nut_allergy", "seafood_allergy", "egg_allergy", "lactose_intolerance", "gluten_intolerance", "soy_allergy", "fish_allergy", "allergy_complete"),
    I14_evs = c("record_id","event_name", "evs_days", "evs_time", "evs_score", "exercise_type_1", "secondary_physical_activity_yn", "tertiary_physical_activity_yn", "evs_complete"),
    I15_alcohol = c("record_id","event_name", "alcohol_frequency", "alcohol_stopped_years_ago", "years_drinks_alcohol", "beer_yn", "wine_yn", "spirit_yn", "alcohol_type_other", "alcohol_type_no_info", "alcohol_type_unknown", "alcohol_type_not_asked", "alcohol_type_asked_unknown", "alcohol_type_invalid", "alcohol_type_na", "beer_per_week", "beer_dose", "wine_ml_per_week", "wine_dose", "distilled_ml_per_week", "spirit_dose", "alcohol_type_other_specify", "alcohol_other_doses_per_week", "alcohol_dose", "alcohol_significant", "alcohol_complete"),
    I16_tobacco = c("record_id","event_name", "smoke_history", "age_started_smoking", "smoked_years", "daily_cigarettes", "pack_years", "age_quit_smoking", "tobacco_complete"),
    I17_diet_recall = c("record_id","event_name", "time_1", "time_2", "time_3", "time_4", "time_5", "time_6", "food_1", "food_2", "food_3", "food_4", "food_5", "food_6", "observations_1", "observations_2", "observations_4", "observations_3", "observations_5", "observations_6", "pdf_upload_recordatorio", "diet_recall_complete", "meal_breakfast_yn", "meal_morning_snack_yn", "meal_lunch_yn", "meal_afternoon_snack_yn", "meal_dinner_yn", "meal_supper_yn", "meal_no_info", "meal_unknown", "meal_not_asked", "meal_asked_unknown", "meal_invalid", "meal_na"),
    I18_diet_values = c("record_id","event_name", "total_energy", "carbs_grams", "carbs_kcal", "protein_kcal", "fat_kcal", "protein_grams", "fat_grams", "carbs_morning_g", "protein_morning_g", "fat_morning_g", "fat_morning_snack_g", "carbs_morning_snack_g", "protein_morning_snack_g", "carbs_lunch_g", "fat_lunch_g", "protein_lunch_g", "carbs_afternoon_snack_g", "fat_afternoon_snack_g", "protein_afternoon_snack_g", "fat_dinner_g", "protein_dinner_g", "carbs_dinner_g", "carbs_night_snack_g", "fat_night_snack_g", "protein_night_snack_g", "carbs_cal_morning", "protein_cal_morning", "fat_cal_morning", "total_cal_morning", "carbs_cal_morning_snack", "fat_cal_morning_snack", "protein_cal_morning_snack", "total_cal_morning_snack", "total_cal_lunch", "carbs_cal_lunch", "fat_cal_lunch", "protein_cal_lunch", "carbs_cal_afternoon_snack", "fat_cal_afternoon_snack", "protein_cal_afternoon_snack", "total_cal_afternoon_snack", "carbs_cal_dinner", "protein_cal_dinner", "fat_cal_dinner", "total_cal_dinner", "carbs_cal_night_snack", "protein_cal_night_snack", "fat_cal_night_snack", "total_cal_night_snack", "nutr_pdf_upload", "diet_values_complete"),
    I19_dates = c("record_id","event_name", "intervention_capsule_start_yn", "visit1_schedule_date", "weekly_alert_remove_yn", "intervention_start_date", "date_45d_complete", "visit2_schedule_date", "date_90d_complete", "visit3_schedule_date", "important_dates_link_hidden", "dates_complete"),
    I20_allocation = c("record_id","event_name", "participant_id", "allocation_group", "allocation_complete"),
    I21_conditions_R = c("record_id","event_name", "repeat_instrument", "repeat_instance", "comorbidity_yn", "comorbidity_additional", "common_comorbidities", "comorbidities_diagnosis_duration", "comorbidities_diagnosis_duration_unit", "comorbidities_listed_concat", "comorbidities_metadata", "conditions_complete"),
    I22_drugs_R = c("record_id","event_name", "repeat_instrument", "repeat_instance", "medication_current_yn", "medication_additional", "common_medications", "medication_dosage", "drug_duration_use", "drug_duration_use_unit", "drug_units_per_dose", "drug_dosage_units", "medication_list_concat", "other_medication_use_yn", "medications_metadata", "drugs_complete"),
    I23_old_drugs_R = c("record_id","event_name", "repeat_instrument", "repeat_instance", "previous_medication_yn", "previous_medication_additional", "common_previous_medications", "previous_drug_start_info", "previous_drug_start_duration", "previous_drug_start_duration_unit", "previous_drug_stop_info", "previous_drug_stop_duration", "previous_drug_stop_duration_unit", "previous_medications_listed_concat", "other_medications", "old.drugs_complete"),
    I24_old_conditions_R = c("record_id","event_name", "repeat_instrument", "repeat_instance", "medical_history_yn", "medical_history_additional", "common_medical_history", "personal_medical_history", "personal_medical_history_notes", "medical_history_duration", "medical_history_duration_unit", "medical_history_listed_concat", "old.conditions_complete"),
    I25_symptoms_R = c("record_id","event_name", "repeat_instrument", "repeat_instance", "symptoms_yn", "symptoms_additional", "symptom_instance_number", "symptom_code", "symptom_specify", "symptom_note", "symptom_onset", "symptom_onset_code", "symptom_onset_unit", "symptom_severity", "symptom_severity_scale", "symptom_current", "symptom_current_or_resolved", "symptom_resolution_date", "symptom_resolution_duration_unit", "symptom_causality_assessment", "symptom_intervention_causality", "symptoms_concat", "symptoms_complete"),
    I26_phy_exam_R = c("record_id","event_name", "repeat_instrument", "repeat_instance", "physical_exam_findings_yn", "physical_exam_findings_additional", "system", "general_appearance", "oral_cavity_exam", "skin_exam", "cardiovascular_exam", "respiratory_exam", "abdominal_exam", "extremities_exam", "abnormal_finding_description", "abnormal_finding_nonindexed_description", "previous_instance_phyexam_concat", "current_instance_phyexam_concat2", "phy.exam_complete"),
    I27_labs_R = c("record_id","event_name", "repeat_instrument", "repeat_instance", "show_lab_list_yn", "lab_request_signed_pdf", "lab_protocol_request_pdf", "medical_consultation_schedule_date", "lab_request_date", "lab_scheduled_date", "lab_location", "labs_date", "c_reactive_protein", "ast", "alt", "gamma_gt", "alkaline_phosphatase", "total_bilirubin", "direct_bilirubin", "indirect_bilirubin", "amylase", "total_proteins", "albumin", "urea", "creatinine", "cpk", "uric_acid", "sodium", "potassium", "calcium_ionic", "total_cholesterol", "ldl", "hba1c", "triglycerides", "hdl", "fasting_glucose", "insulin", "homa_ir", "quick_index", "beta_hcg", "hemoglobin", "hematocrit", "leukocytes", "platelets", "tsh", "t4_free", "labs_notes", "labs_checked_results", "labs_complete"),
    I28_ecg = c("record_id","event_name", "ecg_timestamp", "ecg_performed", "reason_not_performed_ecg", "ecg_pdf_upload", "amplitude_standard", "heart_rate_bpm", "sample_time_s", "qrs_interval_ms", "pr_interval_ms", "qt_interval_ms", "qtc_interval_ms", "qrs_axis_degrees", "rhythm", "normal_ecg", "ecg_findings_lbbb", "ecg_findings_rbbb", "ecg_findings_av_block", "ecg_findings_st_changes", "ecg_findings_abnormal_q_waves", "ecg_findings_nonspecific_intraventricular_conduction_delay", "ecg_findings_no_info", "ecg_findings_unknown", "ecg_findings_not_asked", "ecg_findings_asked_but_unknown", "ecg_findings_invalid", "ecg_findings_na", "ecg_text_prompt", "ecg_audit_data", "ecg_technical_quality", "ecg_measurement_checked", "show_ecg_metadata_yn", "ecg"),
    I29_compliance_R = c("record_id","event_name", "repeat_instrument", "repeat_instance", "taking_as_directed", "dosage_schedule", "alternative_schedule", "reminder_method", "alternative_reminder", "missed_study_medication", "missed_dose_count", "missed_dose_timing", "missed_dose_timing_other", "medication_discontinuation", "discontinuation_duration", "discontinuation_reason", "discontinuation_reason_other", "ran_out_of_medication", "ran_out_reason", "daily_routine_change_medication_adherence_yn", "daily_routine_change_specify", "perceived_improvement", "improvements_reported", "challenges_taking_medication", "specify_challenges_medication_adherence", "medication_confidence_scale", "overall_compliance_rate", "compliance_complete"),
    I30_adverse_event_R = c("record_id","event_name", "repeat_instrument", "repeat_instance", "adverse_event_evaluation_date", "adverse_event_this_cycle_yn", "serious_adverse_event_yn", "adverse_event", "adverse_event_description", "adverse_event_classification", "adverse_event_grade", "adverse_event_start_date", "adverse_event_end_date", "adverse_event_attribution", "adverse_event_action_taken", "adverse_event_followup", "additional_adverse_events_yn", "adverse_event_concat", "adverse_complete"),
    I31_medical = c("record_id","event_name", "evaluation_date", "pipe_comorbidities", "diuretic_use", "diuretic_dose_change", "pipe_current_medications", "pipe_previous_medications", "pipe_medical_history", "menses", "regular_cycle", "cycle_duration", "last_menses", "ovulation_days_after_lmp", "next_ovulation_date", "next_menses_date", "cycle_phase", "drugs_dose_change", "drugs_dose_change_notes", "pipe_symptoms_1st_visit", "pipe_symptoms_2nd_visit", "pipe_symptoms_3rd_visit", "phy_exam_1st_concat", "phy_exam_2nd_concat", "phy_exam_3rd_concat", "show_lab_yn", "lab_tests_yn", "lab_tests_not_performed_reason", "lab_tests_checked_yn", "abnormal_labs", "abnormal_labs_conduct", "intervention_prevention_reason_yn", "specify_intervention_prevention_reasons", "intervention_delivered_yn", "explain_intervention_not_delivered", "date_delivery", "explained_dosage", "intervention_comprehension_yn", "intervention_comprehension_notes", "expected_start_date", "expected_start_date_capsules", "start_after_lab", "additional_notes_actions_yn", "medical_conduct", "medical_complete"),
    I32_followup_R = c("record_id","event_name", "repeat_instrument", "repeat_instance", "weekly_followup_timestamp", "is_data_entry_form_yn", "interviewer_followup", "first_contact_attempt_datetime", "first_contact_successful_yn", "second_contact_attempt_datetime", "second_contact_successful_yn", "third_contact_attempt_datetime", "third_contact_successful_yn", "fourth_contact_attempt_datetime", "fourth_contact_successful_yn", "fifth_contact_attempt_datetime", "fifth_contact_successful_yn", "sixth_contact_attempt_datetime", "sixth_contact_successful_yn", "seventh_contact_attempt_datetime", "seventh_contact_successful_yn", "eighth_contact_attempt_datetime", "eighth_contact_successful_yn", "ninth_contact_attempt_datetime", "ninth_contact_successful_yn", "tenth_contact_attempt_datetime", "tenth_contact_successful_yn", "symptoms_since_treatment", "symptoms_description", "contact_info_start_days", "onset_days", "contact_info_severity", "severity_scale", "symptoms_resolved", "contact_info_duration_days", "resolved_days", "capsules_compliance", "missed_dose_last_week", "reason_not_compliant", "questions_yn", "question_notes", "questions_resolved_yn", "pdf_survey_upload", "followup_complete"),
    I33_conclusion = c("record_id","event_name", "completed_intervention", "conclusion_date", "non_completion_reason", "conclusion_info_view_yn", "ethical_additional_measures_details", "ethical_additional_measures_yn", "ethical_additional_measures_info", "conclusion_complete"),
    I34_annex_R = c("record_id","event_name", "repeat_instrument", "repeat_instance", "attachment_insertion_date", "attachment_type", "attachment_file_date", "attachment_other_specify", "attachment_file_upload", "attachment_notes", "attachment_complete_yn")
)
```
### (?) Saving list to text file
```{r, eval = FALSE}
writeLines(
    capture.output(
        print(instruments)
        ), 
    "/Users/gustavosplmoura/Library/Mobile Documents/com~apple~CloudDocs/Medicina/Biblioteca/Research/Data Science/Data Science/PROJECTS/DVEP/3_Output/4_Instruments.txt")
```

2. Based on the instruments' list, create one tibble for each instrument 
```{r}
# Estas variáveis não devem ser consideradas na verificação de dados faltantes (NA) porque sempre contêm informações.
always_present_vars <- c("record_id", "event_name", "repeat_instrument", "repeat_instance")

# Criar uma lista de tibbles separadas para cada instrumento, excluindo linhas que contenham apenas NAs
instrument_tibbles <- lapply(names(instruments), function(instr_name) {
  
  # `instr_name` é o nome atual do instrumento sendo processado, por exemplo, "elegibility".
  
  # Seleciona a lista de variáveis associadas ao instrumento atual
  selected_vars <- instruments[[instr_name]]
  
  # Remove as variáveis da lista que estão em `always_present_vars` (que sempre possuem valores).
  # `setdiff()` retorna apenas as variáveis exclusivas (aquelas que não estão em `always_present_vars`).
  vars_to_check <- setdiff(selected_vars, always_present_vars)
  
  # Filtrar os dados para o instrumento atual
  filtered_tibble <- data  |> 
    # Seleciona as colunas correspondentes às variáveis do instrumento atual
    select(all_of(selected_vars))  |> 
    
    # Filtra as linhas onde pelo menos uma das variáveis relevantes (não constantes) não é NA
    filter(
      rowSums(
        !is.na(
          pick(all_of(vars_to_check)) # Seleciona apenas as colunas relevantes para a verificação de NA
        )
      ) > 0 # `rowSums()` conta quantas colunas não são NA por linha. Mantemos linhas onde este total é maior que 0.
    )
  
  # Retorna a tibble filtrada com as variáveis e linhas relevantes para o instrumento atual
  return(filtered_tibble)
})

# Nomeia os elementos da lista `instrument_tibbles` com os nomes correspondentes dos instrumentos.
# Por exemplo, o primeiro elemento da lista será nomeado "redcap", o segundo "elegibility", e assim por diante.
names(instrument_tibbles) <- names(instruments)

# Opcional: Salvar cada tibble no ambiente global como um objeto independente.
# `list2env()` converte cada elemento da lista `instrument_tibbles` em um objeto no ambiente global,
# com o nome correspondente ao instrumento.
list2env(instrument_tibbles, .GlobalEnv)

# Resultado:
# - Cada tibble criada contém apenas as variáveis e linhas relevantes para o instrumento atual.
# - As linhas onde todas as variáveis relevantes são NA foram excluídas.

### Corrigindo a lógica de exclusão de linhas com apenas NAs para I20
I20_allocation  <- I20_allocation  |> 
    filter(
        event_name == "eleg_arm_1"
    )

```

## Tibble for non-repeating instruments
List of repeating instruments:
    conditions          comorbidades
    drugs	            medicamentos_de_uso_habitual
    old.drugs           medicamentos_prvios
    old.conditions	    antecedentes_pessoais
    symptoms	        sintomas
    phy.exam	        exame_fsico
    labs	            exames_laboratoriais
    compliance	        adeso
    adverse events	    eventos_adversos
    followup	        contato_semanal
    annex	            anexos

Steps:
1. Filter relevant variables for each visit from `data_codebook`
2. Create tibble for each visit by subsetting `data` with relevant variables
3. Unify data from all visits into a single tibble

Coding:
1. Filter relevant variables for each visit from `data_codebook`
```{r}
## Eleg_arm_1
vars_eleg <- data_codebook  |>
  filter(eleg_arm_1 == 1 & included == 1 & repeating_instrument == 0)  |>
  pull(variable)

## 1visit_arm_1
vars_v1 <- data_codebook  |> 
  filter(V1 == 1 & included == 1 & repeating_instrument == 0)  |>
  pull(variable)

## 2visit_arm_1
vars_v2 <- data_codebook  |> 
  filter(V2 == 1 & included == 1 & repeating_instrument == 0)  |>
  pull(variable)
  
## 3visit_arm_1
vars_v3 <- data_codebook  |> 
  filter(V3 == 1 & included == 1 & repeating_instrument == 0)  |>
  pull(variable)

```

2. Create tibble for each visit by subsetting `data` with relevant variables
```{r}
# First visit (V1)
data_V1 <- data  |> 
  filter(event_name == "eleg_arm_1" & is.na(repeat_instrument)) |> 
  select(setdiff(vars_eleg, c("event_name", "repeat_instrument", "repeat_instance")))  |> 
  mutate(visit = 1) |> 
  relocate(record_id, visit, allocation_group, age, sex)  |> 
  left_join(
    data  |>
      filter(event_name == "1visit_arm_1" & is.na(repeat_instrument)) |> 
      select(setdiff(vars_v1, c("event_name", "repeat_instrument", "repeat_instance"))),
    by = join_by(record_id) # or simply: by = "record_id"
  )  |> # Adding lab results
  left_join(
    I27_labs_R |> 
    filter(event_name == "1visit_arm_1" & repeat_instance == 1)  |> 
    select(record_id, labs_date:labs_notes),
    by = join_by(record_id)
  )

# Second visit (V2)
data_V2 <- data  |> 
  filter(event_name == "2visit_arm_1" & is.na(repeat_instrument)) |> 
  select(setdiff(vars_v2, c("event_name", "repeat_instrument", "repeat_instance")))  |>
  mutate(visit = 2) |>
  relocate(visit, .after = 1
  )|> # Adding lab results
  left_join(
    I27_labs_R |> 
    filter(event_name == "2visit_arm_1" & repeat_instance == 1)  |> 
    select(record_id, labs_date:labs_notes),
    by = join_by(record_id)
  )

# Third visit (V3)
data_V3 <- data  |> 
  filter(event_name == "3visit_arm_1" & is.na(repeat_instrument)) |> 
  select(setdiff(vars_v3, c("event_name", "repeat_instrument", "repeat_instance")))  |> 
  mutate(visit = 3) |> 
  relocate(visit, .after = 1
  )|> # Adding lab results
  left_join(
    I27_labs_R |> 
    filter(event_name == "3visit_arm_1" & repeat_instance == 1)  |> 
    select(record_id, labs_date:labs_notes),
    by = join_by(record_id)
  )
```

3. Unify data from all visits into a single tibble (data_filtered)
```{r}
data_filtered  <- bind_rows(data_V1, data_V2, data_V3)  |> 
  arrange(record_id, visit) |>
  group_by(record_id) |>
  fill(
    allocation_group, age, sex, date_birth, menopause, preserved_hormone, completed_intervention, conclusion_date, non_completion_reason, race, marital_status, education_years, employment_status, household_size, income_level,
  .direction = "down") |>
  ungroup()
```

## Variables grouped by visit/event
```{r}
data_vars_by_visit  <- tibble(
  V1 = c("record_id", "visit", "allocation_group", "age", "sex", "date_birth", "menopause", "preserved_hormone", "completed_intervention", "conclusion_date", "non_completion_reason", "race", "marital_status", "education_years", "employment_status", "household_size", "income_level", "whoqol_score_overall", "whoqol_score_health", "whoqol_score_physical", "whoqol_score_psychological", "whoqol_score_social", "whoqol_score_environment", "dass_score_depression", "dass_score_anxiety", "dass_score_stress", "ecap_score", "height", "weight", "abdomen", "arm", "bmi", "weight_maxideal", "weight_adjusted", "kcal_kg20", "kcal_kg22", "diet_energy", "weight_loss_kg", "weight_loss_percent", "abdomen_loss_kg", "abdomen_loss_percent", "bmi_loss_absolute", "bmi_loss_percent", "systolic_mean", "systolic_stdev", "diastolic_mean", "diastolic_stdev", "mean_bp_mean", "mean_bp_stdev", "smoked_24h", "exercised_24h", "alcohol_24h", "light_clothes", "time_fasted_food", "time_fasted_liquid", "resistance", "reactance", "phase_angle", "bia_complete", "handgrip_right_mean", "handgrip_right_stdev", "handgrip_left_mean", "handgrip_left_stdev", "frequency_bowel", "frequency_solid_stool", "frequency_diarrhea", "contipation_yn", "frequency_laxative", "urinary_frequency", "water_consumption", "hair_loss", "evs_days", "evs_time", "evs_score", "alcohol_frequency", "years_drinks_alcohol", "alcohol_dose", "alcohol_significant", "smoke_history", "age_started_smoking", "smoked_years", "daily_cigarettes", "pack_years", "age_quit_smoking", "total_energy", "carbs_kcal", "protein_kcal", "fat_kcal", "evaluation_date", "diuretic_use", "diuretic_dose_change", "menses", "regular_cycle", "cycle_duration", "last_menses", "cycle_phase", "drugs_dose_change", "drugs_dose_change_notes", "abnormal_labs", "abnormal_labs_conduct", "date_delivery", "explained_dosage", "expected_start_date", "start_after_lab", "medical_conduct"),
  V2 = c("record_id", "visit", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "height", "weight", "abdomen", "arm", "bmi", "weight_maxideal", "weight_adjusted", "kcal_kg20", "kcal_kg22", "diet_energy", "weight_loss_kg", "weight_loss_percent", "abdomen_loss_kg", "abdomen_loss_percent", "bmi_loss_absolute", "bmi_loss_percent", "systolic_mean", "systolic_stdev", "diastolic_mean", "diastolic_stdev", "mean_bp_mean", "mean_bp_stdev", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "frequency_bowel", "frequency_solid_stool", "frequency_diarrhea", "contipation_yn", "frequency_laxative", "urinary_frequency", "water_consumption", "hair_loss", "evs_days", "evs_time", "evs_score", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "evaluation_date", "diuretic_use", "diuretic_dose_change", "menses", "regular_cycle", "cycle_duration", "last_menses", "cycle_phase", "drugs_dose_change", "drugs_dose_change_notes", "abnormal_labs", "abnormal_labs_conduct", "date_delivery", "explained_dosage", "expected_start_date", "start_after_lab", "medical_conduct"),
  V3 = c("record_id", "visit", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "whoqol_score_overall", "whoqol_score_health", "whoqol_score_physical", "whoqol_score_psychological", "whoqol_score_social", "whoqol_score_environment", "dass_score_depression", "dass_score_anxiety", "dass_score_stress", "ecap_score", "height", "weight", "abdomen", "arm", "bmi", "weight_maxideal", "weight_adjusted", "kcal_kg20", "kcal_kg22", "diet_energy", "weight_loss_kg", "weight_loss_percent", "abdomen_loss_kg", "abdomen_loss_percent", "bmi_loss_absolute", "bmi_loss_percent", "systolic_mean", "systolic_stdev", "diastolic_mean", "diastolic_stdev", "mean_bp_mean", "mean_bp_stdev", "smoked_24h", "exercised_24h", "alcohol_24h", "light_clothes", "time_fasted_food", "time_fasted_liquid", "resistance", "reactance", "phase_angle", "bia_complete", "handgrip_right_mean", "handgrip_right_stdev", "handgrip_left_mean", "handgrip_left_stdev", "frequency_bowel", "frequency_solid_stool", "frequency_diarrhea", "contipation_yn", "frequency_laxative", "urinary_frequency", "water_consumption", "hair_loss", "evs_days", "evs_time", "evs_score", "alcohol_frequency", "years_drinks_alcohol", "alcohol_dose", "alcohol_significant", "smoke_history", "age_started_smoking", "smoked_years", "daily_cigarettes", "pack_years", "age_quit_smoking", "total_energy", "carbs_kcal", "protein_kcal", "fat_kcal", "evaluation_date", "diuretic_use", "diuretic_dose_change", "menses", "regular_cycle", "cycle_duration", "last_menses", "cycle_phase", "drugs_dose_change", "drugs_dose_change_notes", "abnormal_labs", "abnormal_labs_conduct", "date_delivery", "explained_dosage", "expected_start_date", "start_after_lab", "medical_conduct") 
)
```

### Assign labels to data_filtered
```{r}
# Load Hmisc for `label()` function
library(Hmisc)

# Get the variable names from the data_filtered dataframe
variable_names <- colnames(data_filtered)

# Loop over each variable name in data_filtered
for (variable_name in variable_names) {
    # Check if the variable name exists in the data_codebook$variable
    if (variable_name %in% data_codebook$variable) {
        # Get the corresponding label for the variable
        variable_label <- data_codebook$label_pt[data_codebook$variable == variable_name]
        
        # Assign the label to the variable in data_filtered
        label(data_filtered[[variable_name]]) <- variable_label
    }
}

# Remove intermediate variables
rm(variable_name, variable_label)
```

#### Save data_filtered as csv
```{r}
write_csv(data_filtered, "/Users/gustavosplmoura/Library/Mobile Documents/com~apple~CloudDocs/Medicina/Biblioteca/Research/Data Science/Data Science/PROJECTS/DVEP/data_filtered.csv", na = "")
```

## Adding labels to raw cell values
To apply the labels only to the variables present in colnames(data_filtered), you can filter the data_codebook to include only those variables.
Steps for Selected Variables
1. Filter data_codebook for variables present in colnames(data_filtered).
2. Parse the choices column for these variables only.
3. Create lookup tables for the selected variables.
4. Apply the mappings to replace raw values with their corresponding labels in data_filtered.

```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)

# 1. Filter `data_codebook` for variables in `data_filtered` and with `col_types` == "f"
selected_codebook <- data_codebook  |> 
  filter(
    variable %in% colnames(data_filtered) &  # Variables in data_filtered
      col_types == "f"                              # Only those with col_types == "f"
  )

# 2. Parse the `choices` column
parsed_choices <- selected_codebook |>
  rowwise() |>
  mutate(
    parsed = list(
      str_split(choices, " \\| ") |>  # Split the choices by "|"
        unlist() |>
        map(~ str_split_fixed(.x, ", ", 2) |>  # Split by ", " into two columns
          as_tibble()  |> 
          setNames(c("raw_value", "label"))    # Explicitly name columns
        ) |>
        bind_rows()  # Combine into a tibble
    )
  ) |>
  select(variable, parsed) |>
  unnest(parsed)  # Expand the parsed choices into rows


# 3. Create lookup tables for selected variables
lookup_tables <- parsed_choices %>%
  group_by(variable) %>%
  summarize(
    lookup = list(setNames(label, raw_value)), .groups = "drop"
  ) %>%
  deframe()

# 4. Replace raw values with labels for selected variables

data_filtered_labels <- data_filtered
for (column_name in names(lookup_tables)) { # Iterate through each column in lookup_tables
  if (column_name %in% colnames(data_filtered_labels)) { # Ensure the column exists in data_filtered
    data_filtered_labels[[column_name]] <- recode( # Apply recode using the lookup table for this column
      data_filtered_labels[[column_name]],
      !!!lookup_tables[[column_name]]
    )
  }
}
```

## Repeating instruments
Steps:
1. Concatenating NCIT labels (from data_NCIT) to NCIT codes in data
```{r}
# 1.1 Conditions / commorbidities
I21_conditions_R  <- I21_conditions_R  |> 
    left_join(
        data_NCIT |> select(ncit_code, descriptive),
        # by = c("common_comorbidities" = "ncit_code")
        join_by(common_comorbidities == ncit_code)
    ) |>
    relocate(
        descriptive, .after = common_comorbidities
    )

# 1.2 Drugs in regular use
I22_drugs_R  <- I22_drugs_R  |> 
    left_join(
        data_NCIT |> select(ncit_code, descriptive),
        join_by(common_medications == ncit_code)
    ) |>
    relocate(
        descriptive, .after = common_medications
    )

# 1.3 Previous drugs
I23_old_drugs_R <- I23_old_drugs_R |> 
    left_join(
        data_NCIT |> select(ncit_code, descriptive),
        join_by(common_previous_medications == ncit_code)
    )|>
    relocate(
        descriptive, .after = common_previous_medications
    )

# 1.4 Past medical conditions
I24_old_conditions_R <- I24_old_conditions_R |> 
    left_join(
        data_NCIT |> select(ncit_code, descriptive),
        join_by(common_medical_history == ncit_code)
    )|>
    relocate(
        descriptive, .after = common_medical_history
    )
```

2. Most common comorbidities
```{r}
I21_conditions_R |> 
    group_by(common_comorbidities, descriptive) |> 
    count(common_comorbidities, sort = TRUE, name = "frequency") |> 
    mutate(percentage = frequency/74 * 100)
```
```{r}
data_V1 <- data_V1 |> 
    mutate(
        hypertension = 
            if_else(
                record_id %in% (I21_conditions_R |> 
                                     filter(str_detect(descriptive, "Hipertensão")) |> 
                                     pull(record_id)),
                1, 0
                )
        ) |> 
    relocate(hypertension, .after = income_level)

```




