

```{r}
library(tidyverse)

# Function to create a summary table
create_summary_table <- function(data, variables = NULL) {
  # If no variables specified, use all numeric columns
  if (is.null(variables)) {
    variables <- names(data)[sapply(data, is.numeric)]
  }
  
  # Check if specified variables exist in the dataframe
  valid_variables <- intersect(variables, colnames(data))
  
  # If no valid variables are found, return an error
  if (length(valid_variables) == 0) {
    stop("No valid numeric variables found in the provided dataframe.")
  }
  
  # Create the summary table
  summary_table <- data %>%
    summarise(across(
      all_of(valid_variables),
      list(
        mean = ~ mean(.x, na.rm = TRUE),
        sd = ~ sd(.x, na.rm = TRUE),
        median = ~ median(.x, na.rm = TRUE),
        iqr = ~ IQR(.x, na.rm = TRUE),
        min = ~ min(.x, na.rm = TRUE),
        max = ~ max(.x, na.rm = TRUE),
        missing = ~ sum(is.na(.x))
      )
    )) %>%
    pivot_longer(
      cols = everything(),
      names_to = c("variable", "statistic"),
      names_sep = "_",
      values_to = "value"
    ) %>%
    pivot_wider(
      names_from = statistic,
      values_from = value
    )
  
  return(summary_table)
}

```

```{r}
library(lme4)

glmm_model <- lme4::lmer(
  phaseangle ~ visit + allocation_group + age + sex + labs_homa_ir + labs_leukocytes +
    (1 | record_id),  # Random effect for patients
  data = final_data,
  REML = FALSE        # Use maximum likelihood for comparison between models
)

# View the model summary
summary(glmm_model)


glmm_model_interaction <- lmer(
  phaseangle ~ visit * allocation_group + age + sex + labs_homa_ir + labs_leukocytes +
    (1 | record_id),
  data = final_data
)
summary(glmm_model_interaction)


```

```{r}
# Fit the linear mixed-effects model for CRP
crp_model <- lmer(
  labs_crp ~ visit * allocation_group + age + sex + labs_homa_ir + labs_leukocytes +
    (1 | record_id),  # Random intercept for repeated measures
  data = final_data
)

# View the summary
summary(crp_model)

```

# Help

## Count()

Used for categorical data

```{r, eval=FALSE}
# Count Unique Values in a Single Column
d1_exclusive |> 
  count(allocation_group)

# Count Unique Combinations of Two or More Variables
d1_exclusive |> 
  count(allocation_group, visit)

#  Add Proportions Using count()
d1_exclusive |> 
  count(allocation_group, name = "count") |> 
  mutate(proportion = count / sum(count))

# Sort the Results
d1_exclusive |> 
  count(allocation_group, sort = TRUE)

# Filter for Specific Counts
d1_exclusive |> 
  count(allocation_group) |> 
  filter(n > 20)

# Rename the Count Column
d1_exclusive |> 
  count(allocation_group, name = "total")

# Weight Counts
You can apply weights to the counts (e.g., if certain rows represent different weights)

d1_exclusive |> 
  count(allocation_group, wt = weight_column)

## Here, weight_column provides the weights to adjust the counts. If weight_column = c(2, 1, 1, ...), the count for Grupo A might become 37 × 2.

# Use count() in Pipelines
d1_exclusive |> 
  group_by(sex) |> 
  count(allocation_group) |> 
  arrange(desc(n))

# Visualize the Counts
d1_exclusive |> 
  count(allocation_group) |> 
  ggplot(aes(x = allocation_group, y = n, fill = allocation_group)) +
  geom_bar(stat = "identity") +
  labs(title = "Participant Count by Group", x = "Group", y = "Count") +
  theme_minimal()



```

# Jamovi

## jmv::descriptives

Defaults

```{r}
library(jmv)
#Default
# N, missing, mean, median, SD, Min, Max
jmv::descriptives(
    data = data, # select dataframe
    vars = vars() #insert variables without ""
    ) 
```

To remove defaults

```{r}
jmv::descriptives(
    data = data,
    vars = vars(),
    n = FALSE,
    missing = FALSE,
    mean = FALSE,
    median = FALSE,
    sd = FALSE,
    min = FALSE,
    max = FALSE
    )
```

Additional statistics

```{r}

```

jmv::descriptives( data = data, vars = vars(), pcEqGr = TRUE, pcNEqGr = 5, pc = TRUE)

jmv::descriptives( data = data, vars = vars(age, abdomen, bmi.x), freq = TRUE, desc = "rows", boxLabelOutliers = FALSE, mode = TRUE, sum = TRUE, variance = TRUE, range = TRUE, se = TRUE, ci = TRUE, iqr = TRUE, skew = TRUE, kurt = TRUE, sw = TRUE, pcEqGr = TRUE, pc = TRUE, extreme = TRUE)

```{r}

# Extract vector of double variables
dbl_vars <- final_data %>%
  select(where(is.double)) %>%
  colnames()

# View the result
print(dbl_vars)

jmv::descriptives(
    data = final_data,
    vars = !!rlang::syms(dbl_vars),
    splitBy = NULL,
    freq = FALSE, #frequency tables (nominal, ordinal variables only)
    desc = "columns", # 'rows' or 'columns' to display variables
    hist = FALSE, # histograms (continuous variables only)
    dens = FALSE, # density plots (continuous variables only)
    bar = FALSE, # bar plots (nominal, ordinal variables only)
    barCounts = FALSE, # add counts to the bar plots
    box = FALSE, # box plots (continuous variables only)
    violin = FALSE, # violin plots (continuous variables only)
    dot = FALSE, # dot plots (continuous variables only)
    dotType = "jitter",
    boxMean = FALSE, # add mean to box plot
    boxLabelOutliers = TRUE, # labels outliers with the row
    qq = FALSE, # (continuous variables only)
    n = TRUE, # sample size
    missing = TRUE, # number of missing values
    mean = TRUE,
    median = TRUE,
    mode = FALSE,
    sum = FALSE,
    sd = TRUE,
    variance = FALSE,
    range = FALSE,
    min = TRUE,
    max = TRUE,
    se = FALSE, # standard error of the mean
    ci = FALSE, # confidence intervals for the mean
    ciWidth = 95, # width of confidence intervals (50-99.9)
    iqr = FALSE,
    skew = FALSE,
    kurt = FALSE,
    sw = FALSE, # Shapiro-Wilk p-value
    pcEqGr = FALSE, # provide quantiles (default: 4)
    pcNEqGr = 4, # an integer specifying the number of equal groups
    pc = FALSE, # provide percentiles
    pcValues = "25,50,75", #specify the percentiles
    extreme = FALSE, # provide N most extreme (highest/lowest) values
    extremeN = 5, # specify the number of extreme values
    )
```
