---
title: "seca_bioimpedance_data"
output: html_document
---

```{r}
codebook_bia <- read_excel(
    "Codebooks/codebook_bia.xlsx",
    col_names = TRUE,
    col_types = NULL,
        na = c("", "NA", "NI", "UNK", "NASK", "ASKU", "INV"),
        trim_ws = TRUE,
        skip = 0, # Number of lines to skip before reading data
        n_max = Inf, # Maximum number of lines to read.
        guess_max = 1000
    ) |>
    arrange(index)

data_bia_D3 <- read_csv(
    "Data/data_bia_D3.csv",
    col_names = TRUE)

data_bia_D1 <- read_csv(
    "Data/data_bia_D1.csv",
    col_names = TRUE)
```

## record_id as.integer

```{r}
# 5. record_id as.integer
data_bia_D3 $record_id <- as.integer(data_bia_D3 $record_id)
data_bia_D1$record_id <- as.integer(data_bia_D1$record_id)
```

## Renaming variables
```{r}
# 4.2 bia
rename_bia <- setNames(object = colnames(data_bia_D3 ), codebook_bia$variable)

data_bia_D3 <- data_bia_D3 |>
    rename(!!!rename_bia)

data_bia_D1 <- data_bia_D1 |>
    rename(!!!rename_bia)

rm(rename_bia)
```

## Assign labels to variables

```{r}
data_bia_D3 <- data_bia_D3 |> 
  mutate(across(
    all_of(codebook_bia$variable),
    ~ {
      attr(., "label") <- codebook_bia$label_pt[codebook_bia$variable == cur_column()]
      .
    }
  ))
```

## Bioimpedance data (DIRECT IMPORT FROM BIOIMPEDANCE DEVICE)

This corresponds to the first version of the code. However, it relies on data directly imported from the bioimpedance device, and contains one to three measurements per participant. Therefore, the data needs to be wrangled to obtain the mean of the measurements for each participant.

However, for now (April 09th, 2025), I will depracate this code and analyse the data directly recorded into the REDCap project by the research collaborators. The code below is kept for reference, but will not be used in the analysis as of now. If needed, I will re-implement it in the future.

#### Data from first/third visit

Applies to participants who completed the intervention

```{r, eval=FALSE, echo=FALSE}
# 8. Wrangling Bioimpedance data
## 8.1 D3 DATA (contains data from first and third visits for participants who completed the intervention)
### Filter lines for which phaseangle is <> ""
data_bia_D3_filtered <- data_bia_D3 |> 
    filter(!is.na(phaseangle)) |> 
    mutate(
        date = as.Date(timestamp),           # Extract the date
        time = format(timestamp, "%H:%M:%S") # Extract the time
    ) |> 
    select(all_of(codebook_bia$variable)[codebook_bia$included == 1], date, time) |> 
    relocate(record_id, date, time, phaseangle, raverage, xcaverage, weight:w_ecwbytbw) |> 
    arrange(record_id, date, time)

### Group by record_id and date and obtain mean of multiple measurements from the same day
data_bia_D3_filtered <- data_bia_D3_filtered |> 
    group_by(record_id, date) |>
    summarise(
        across(c(phaseangle:m_tohimaginary), \(x) mean(x, na.rm = TRUE)),
        .groups = "drop"
    ) |>
    group_by(record_id) |> # Add coding for visit number
    mutate(
        visit = case_when(
            date == min(date) ~ 1,  # Assign 1 to the earliest date
            date == max(date) ~ 3,  # Assign 3 to the latest date
            TRUE ~ NA_real_         # Default to NA for unexpected cases
        ),
        .after = record_id
    )
```

#### Data from first visit

Applies to participants who did not complete the intervention

```{r, eval=FALSE, echo=FALSE}
## 8.2 D1 DATA (data from the first visit for participants who did not complete the intervention)
## Filter lines for which phaseangle is <> ""
data_bia_D1_filtered <- data_bia_D1 |> 
    filter(!is.na(phaseangle)) |> 
    mutate(
        date = as.Date(timestamp),           # Extract the date
        time = format(timestamp, "%H:%M:%S") # Extract the time
    ) |> 
    select(all_of(codebook_bia$variable)[codebook_bia$included == 1], date, time) |> 
    relocate(record_id, date, time, phaseangle, raverage, xcaverage, weight:w_ecwbytbw) |> 
    arrange(record_id, date, time)

### Group by record_id and date and obtain mean of multiple measurements from the same day
data_bia_D1_filtered <- data_bia_D1_filtered |> 
    group_by(record_id, date) |>
    summarise(
        across(c(phaseangle:m_tohimaginary), \(x) mean(x, na.rm = TRUE)),
        .groups = "drop"
    ) |>
    group_by(record_id) |> # Add coding for visit number
    mutate(
        visit = case_when(
            date == min(date) ~ 1,  # Assign 1 to the earliest date
            date == max(date) ~ 3,  # Assign 3 to the latest date
            TRUE ~ NA_real_         # Default to NA for unexpected cases
        ),
        .after = record_id
    )
### Selecting BIA data from D1 not present in D3
data_bia_D1_filtered <- data_bia_D1_filtered |> 
    filter(
     record_id %in% setdiff(1:75, data_bia_D3_filtered$record_id)   
    )
```

#### Merging to single tibble

```{r, eval=FALSE, echo=FALSE}
## 8.3 MERGE D1/D3 BIA data into a single tibble
data_bia <- bind_rows(
    data_bia_D1_filtered, data_bia_D3_filtered
) |> 
    mutate(
        visit = as.integer(visit)
    ) |> 
    arrange(
        record_id, visit
    ) |> 
    ungroup()

### label_variables
data_bia <- label_variables(data_bia, codebook_bia)
```

#### Drop intermediate tibbles

```{r, eval=FALSE, echo=FALSE}
# 8.4. DROP intermediate tibbles
rm(data_bia_D1)
rm(data_bia_D1_filtered)
rm(data_bia_D3)
rm(data_bia_D3_filtered)
```
