---
title: "whoqol-bref"
format: html
---

# SCORING OF THE WHOQOL-BREF

The WHOQOL-BREF (Field Trial Version) produces four domain scores. There are also two items that are examined separately: question 1 asks about an individual’s overall perception of quality of life and question 2 asks about an individual’s overall perception of his or her health.

Domain scores are scaled in a positive direction (i.e., higher scores denote higher quality of life). The mean score of items within each domain is used to calculate the domain score. Mean scores are then multiplied by 4 in order to make domain scores comparable with the scores used in the WHOQOL-100, and subsequently transformed to a 0–100 scale, using the formula above.

Explicit instructions for checking and cleaning data, and for computing domain scores, are given in Appendix 10. A method for the manual calculation of individual scores is below:

### Domain Scoring Formulas

-   **Physical domain** = `((6 - Q3) + (6 - Q4) + Q10 + Q15 + Q16 + Q17 + Q18) × 4`
-   **Psychological domain** = `(Q5 + Q6 + Q7 + Q11 + Q19 + (6 - Q26)) × 4`
-   **Social Relationships domain** = `(Q20 + Q21 + Q22) × 4`
-   **Environment domain** = `(Q8 + Q9 + Q12 + Q13 + Q14 + Q23 + Q24 + Q25) × 4`

> Note: The values `(6 - Q3)`, `(6 - Q4)`, and `(6 - Q26)` correspond to the recoding of negatively phrased items into positively phrased scales.

### Missing Data Rules

-   Where more than **20% of data** are missing from an assessment, the assessment should be discarded (see Step 4 in Table 3).
-   Where up to two items are missing, the **mean of other items in the domain is substituted**.
-   Where more than two items are missing from the domain, the **domain score should not be calculated** (with the exception of domain 3, where the domain should only be calculated if ≤ 1 item is missing).

> Syntax files for automatic computation of domains are available as for the WHOQOL-100.

## Appendix 10 – Steps for Checking and Cleaning Data and Computing Domain Scores for the WHOQOL-BREF

### 1. Check all 26 items from assessment have a range of 1–5

```spss
RECODE Q1 Q2 Q3 Q4 Q5 Q6 Q7 Q8 Q9 Q10 Q11 Q12 Q13 Q14 Q15 Q16 Q17 Q18 Q19 Q20 Q21 Q22 Q23 Q24 Q25 Q26 (1=1) (2=2) (3=3) (4=4) (5=5) (ELSE=SYSMIS).
```

*This recodes all data outside the range 1–5 to system missing.*

---

### 2. Reverse 3 negatively phrased items

```spss
RECODE Q3 Q4 Q26 (1=5) (2=4) (3=3) (4=2) (5=1).
```

*This transforms negatively framed questions to positively framed questions.*

---

### 3. Compute domain scores

```spss
COMPUTE PHYS = MEAN.6(Q3, Q4, Q10, Q15, Q16, Q17, Q18) * 4.
COMPUTE PSYCH = MEAN.5(Q5, Q6, Q7, Q11, Q19, Q26) * 4.
COMPUTE SOCIAL = MEAN.2(Q20, Q21, Q22) * 4.
COMPUTE ENVIR = MEAN.6(Q8, Q9, Q12, Q13, Q14, Q23, Q24, Q25) * 4.
```

*These equations calculate the domain scores. All scores are multiplied by 4 to make them directly comparable with scores from the WHOQOL-100. The number after `MEAN.` (e.g., 6) specifies the minimum number of items that must be present to compute the domain.*

---

### 4. Transform scores to a 0–100 scale

```spss
COMPUTE PHYS = (PHYS - 4) * (100 / 16).
COMPUTE PSYCH = (PSYCH - 4) * (100 / 16).
COMPUTE SOCIAL = (SOCIAL - 4) * (100 / 16).
COMPUTE ENVIR = (ENVIR - 4) * (100 / 16).
```

---

### 5. Delete cases with >20% missing data

```spss
COUNT TOTAL = Q1 TO Q26 (1 THRU 5).
SELECT IF (TOTAL >= 21).
EXECUTE.
```

*This command creates a new column `TOTAL` that counts valid responses (1–5) for Q1 to Q26. If fewer than 21 responses (i.e., <80%) are available, the case is excluded.*


## WHOQOL-BREF Scoring Formulas (REDCap Logic)

The following are the calculation formulas used to score the WHOQOL-BREF domains and individual perception items, implemented using `@CALCTEXT` in REDCap. All domain scores are transformed to a 0–100 scale.

### Overall Quality of Life

```text
whoqol_score_overall
@CALCTEXT(if([wb_1] = '', '', (([wb_1:value] * 4) - 4) * (100 / 16)))
@HIDDEN
```

### Overall Perception of Health

```text
whoqol_score_health
@CALCTEXT(if([wb_2] = '', '', (([wb_2:value] * 4) - 4) * (100 / 16)))
@HIDDEN
```

### Physical Health Domain

```text
whoqol_score_physical
@CALCTEXT(
  if([wb_3] = '' or [wb_4] = '' or [wb_10] = '' or [wb_15] = '' or [wb_16] = '' or [wb_17] = '' or [wb_18] = '',
     '',
     ((((sum((6 - [wb_3:value]), (6 - [wb_4:value]), [wb_10:value], [wb_15:value], [wb_16:value], [wb_17:value], [wb_18:value])) * 4) / 7) - 4) * (100 / 16)
  )
)
@HIDDEN
```

### Psychological Domain

```text
whoqol_score_psychological
@CALCTEXT(
  if([wb_5] = '' or [wb_6] = '' or [wb_7] = '' or [wb_11] = '' or [wb_19] = '' or [wb_26] = '',
     '',
     (((sum([wb_5:value], [wb_6:value], [wb_7:value], [wb_11:value], [wb_19:value], (6 - [wb_26:value])) * 4) / 6) - 4) * (100 / 16)
  )
)
@HIDDEN
```

### Social Relationships Domain

```text
whoqol_score_social
@CALCTEXT(
  if([wb_20] = '' or [wb_21] = '' or [wb_22] = '',
     '',
     ((((sum([wb_20:value], [wb_21:value], [wb_22:value])) * 4) / 3) - 4) * (100 / 16)
  )
)
@HIDDEN
```

### Environment Domain

```text
whoqol_score_environment
@CALCTEXT(
  if([wb_8] = '' or [wb_9] = '' or [wb_12] = '' or [wb_13] = '' or [wb_14] = '' or [wb_23] = '' or [wb_24] = '' or [wb_25] = '',
     '',
     ((((sum([wb_8:value], [wb_9:value], [wb_12:value], [wb_13:value], [wb_14:value], [wb_23:value], [wb_24:value], [wb_25:value])) * 4) / 8) - 4) * (100 / 16)
  )
)
@HIDDEN
```

### File Upload Field

```text
pdf_upload_wb
@HIDDEN-SURVEY  
@HIDDEN-PDF
```

Let me know if you want this formatted into a downloadable `.qmd` file or used to create a REDCap data dictionary or report.

# LOAD DATA

```{r}
library(tidyverse)
data <- readRDS("local_files/Data_processed/data.rds")
whoqol <- data %>% 
    select(whoqol_timestamp:whoqol_score_environment) %>% 
    filter(
        !is.na(whoqol_timestamp)
    )

names(whoqol)
```

```{r}
# Reverse-score the 3 negatively phrased items: Q3, Q4, Q26
whoqol$whoqol_3_pain       <- 6 - whoqol$whoqol_3_pain       # Q3
whoqol$whoqol_4_treatment  <- 6 - whoqol$whoqol_4_treatment  # Q4
whoqol$whoqol_26_negativity <- 6 - whoqol$whoqol_26_negativity # Q26

# Physical Health Domain: Q3, Q4, Q10, Q15, Q16, Q17, Q18
whoqol$score_physical <- rowMeans(whoqol[, c(
  "whoqol_3_pain", "whoqol_4_treatment", "whoqol_10_energy", 
  "whoqol_15_mobility", "whoqol_16_sleep", "whoqol_17_activities", 
  "whoqol_18_work"
)], na.rm = TRUE) * 4

# Psychological Domain: Q5, Q6, Q7, Q11, Q19, Q26
whoqol$score_psychological <- rowMeans(whoqol[, c(
  "whoqol_5_enjoyment", "whoqol_6_meaning", "whoqol_7_concentration",
  "whoqol_11_appearance", "whoqol_19_selfesteem", "whoqol_26_negativity"
)], na.rm = TRUE) * 4

# Social Relationships Domain: Q20, Q21, Q22
whoqol$score_social <- rowMeans(whoqol[, c(
  "whoqol_20_relationships", "whoqol_21_sexual", "whoqol_22_support"
)], na.rm = TRUE) * 4

# Environment Domain: Q8, Q9, Q12, Q13, Q14, Q23, Q24, Q25
whoqol$score_environment <- rowMeans(whoqol[, c(
  "whoqol_8_security", "whoqol_9_environment", "whoqol_12_finances",
  "whoqol_13_information", "whoqol_14_leisure", "whoqol_23_housing",
  "whoqol_24_health_services", "whoqol_25_transport"
)], na.rm = TRUE) * 4

# Overall Perception of Quality of Life (Q1)
whoqol$score_overall_quality <- (whoqol$whoqol_1_quality * 4 - 4) * (100 / 16)

# Overall Perception of Health (Q2)
whoqol$score_overall_health <- (whoqol$whoqol_2_health * 4 - 4) * (100 / 16)

# Transform domain scores to 0–100 scale
whoqol$score_physical      <- (whoqol$score_physical - 4) * (100 / 16)
whoqol$score_psychological <- (whoqol$score_psychological - 4) * (100 / 16)
whoqol$score_social        <- (whoqol$score_social - 4) * (100 / 16)
whoqol$score_environment   <- (whoqol$score_environment - 4) * (100 / 16)
```

```{r}
# Helper function: compute row mean only if at least N non-NA values
mean_if_enough <- function(x, min_valid) {
  if (sum(!is.na(x)) >= min_valid) {
    return(mean(x, na.rm = TRUE))
  } else {
    return(NA_real_)
  }
}
```

```{r}
# Physical domain: require at least 6 of 7
whoqol$score_physical <- apply(whoqol[, c(
  "whoqol_3_pain", "whoqol_4_treatment", "whoqol_10_energy", 
  "whoqol_15_mobility", "whoqol_16_sleep", "whoqol_17_activities", 
  "whoqol_18_work"
)], 1, mean_if_enough, min_valid = 6) * 4

# Psychological domain: require at least 5 of 6
whoqol$score_psychological <- apply(whoqol[, c(
  "whoqol_5_enjoyment", "whoqol_6_meaning", "whoqol_7_concentration",
  "whoqol_11_appearance", "whoqol_19_selfesteem", "whoqol_26_negativity"
)], 1, mean_if_enough, min_valid = 5) * 4

# Social domain: require at least 2 of 3
whoqol$score_social <- apply(whoqol[, c(
  "whoqol_20_relationships", "whoqol_21_sexual", "whoqol_22_support"
)], 1, mean_if_enough, min_valid = 2) * 4

# Environment domain: require at least 6 of 8
whoqol$score_environment <- apply(whoqol[, c(
  "whoqol_8_security", "whoqol_9_environment", "whoqol_12_finances",
  "whoqol_13_information", "whoqol_14_leisure", "whoqol_23_housing",
  "whoqol_24_health_services", "whoqol_25_transport"
)], 1, mean_if_enough, min_valid = 6) * 4
```


```{r}
whoqol$score_physical      <- (whoqol$score_physical - 4) * (100 / 16)
whoqol$score_psychological <- (whoqol$score_psychological - 4) * (100 / 16)
whoqol$score_social        <- (whoqol$score_social - 4) * (100 / 16)
whoqol$score_environment   <- (whoqol$score_environment - 4) * (100 / 16)

```

