Yes, you can use an R script to run another R script. There are several ways to do this, depending on your specific needs. Here are a few common approaches:

### 1. Using `source()` function
The `source()` function can be used to run an R script from within another R script. The `source()` function reads and evaluates the expressions in the specified file.

Example:
```{r}
# Main script
print("Running the main script")

# Run another R script
source("path/to/your_script.R")

print("Finished running the other script")
```

In this example, `"path/to/your_script.R"` is the file path to the script you want to run. You can also pass additional arguments, such as `local = TRUE` if you want the variables created in the sourced script to only exist in the local environment.

### 2. Using `system()` or `system2()` to call the Rscript command
If you want to run another R script as a separate process, you can use the `system()` or `system2()` functions to call the Rscript command-line utility. This method runs the script in a new R session.

Example using `system()`:
```{r}
# Run another R script using the system command
system("Rscript path/to/your_script.R")
```

Example using `system2()`:
```{r}
# Run another R script using the system2 command
system2("Rscript", args = "path/to/your_script.R")
```

### 3. Using `source()` with arguments
If you want to pass arguments to the script you're running, you can handle arguments in the script using `commandArgs()` in the R script being sourced.

Example in the main script:
```{r}
source("path/to/your_script_with_args.R")
```

In `your_script_with_args.R`, use `commandArgs()` to capture the arguments:
```{r}
args <- commandArgs(trailingOnly = TRUE)
print(args)
```

### 4. Scheduling scripts using `taskscheduleR` (Windows) or cron (Linux/macOS)
While this isn't directly within an R script, you can also schedule scripts using task schedulers like `taskscheduleR` on Windows or cron jobs on Linux/macOS to run multiple R scripts automatically at specified times.

### Conclusion
Yes, it's absolutely possible to run another R script from within an R script using any of these methods, depending on how you'd like the second script to be executed and what environment you want it to run in. The most common approach is using the `source()` function for simplicity, but running the script as a separate process with `system()` or `system2()` might be more appropriate in some situations.